### 2020-08-29 Sections/costs_and_attributes_in_quantum_networks.tex
editing Saturday night

The fidelity of two states directly quantifies how close they are to one another in a geometric sense, i.e on the Bloch sphere \cite{???}, or, in the context of a state passing through a quantum channel, a measure of how well the state is preserved.

@@

%
% Channel Capacity
%

\subsubsection{Channel capacity} \label{sec:channel_cap} \index{Channels!Capacity}

The measures considered until now have quantified the preservation of quantum states. Alternately, one might consider information theoretic measures\index{Information theoretic!Measures}, which quantify the number of bits/qubits transmitted by a link, i.e the number of bits/qubits in common before and after the channel. This is an extremely powerful tool as it upper bounds the amount of information the receiver can extract from the transmitter under \textit{any} measurement scheme, very useful in a cryptographic context, where we want security to be attack-independent\index{Information-theoretic!Security} (Sec.~\ref{sec:comp_vs_inf_th_sec}). Before we delve into this, we have to introduce a few new quantities.

\paragraph{Classical correlation measures}

Suppose we have a random variable $X$, and a particular measurement outcome $x$ occurs with probability $p(x)$, the Shannon entropy of $X$ is given by,
\begin{align}\index{Shannon entropy}
H(X) = -\sum_x p_x\log_2(p_x).
\end{align}

The information, or entropy, plays a central role in information theory. The intuitive interpretation is that it quantifies an experimenter's uncertainty about $X$ before measuring it, and his expected information gain is $H(X)$ bits upon learning the outcome. The information $H(X)$ is zero if and only if one of the probabilities $p(x)$ is unity, with the others being zero. In this case the value of $X$ is already known and so there is no information to be gained from observing it.

 For a joint distribution over $X$ and $Y$ this simply generalises to,
\begin{align}\index{Joint Shannon entropy}
H(X,Y) =  -\sum_{x,y} p_{x,y}\log_2(p_{x,y}).
\end{align}

The von Neuman entropy\index{von Neuman entropy} \cite{bib:bengtsson2017geometry} for quantum density operators, $S(\hat\rho)$, is defined analogously, replacing probabilities with density operator eigenvalues,
\begin{align}\index{von Neuman entropy}
S(\hat\rho) &= - \sum_x \lambda_x \log_2 (\lambda_x) \nonumber \\
&= -\mathrm{tr}(\hat\rho\,\log \,\hat\rho),
\end{align}
where $\{\lambda\}$ is the eigenvalue spectrum of $\hat\rho$. This modification is logically justified, as the eigenvalues can be interpreted directly as a purely classical probability distribution of orthogonal states when the density operator is transformed into a basis with no coherences between basis states (i.e a diagonal basis or spectral decomposition\index{Spectral decomposition}). In that case, the Shannon and von Neuman entropies essentially have identical physical interpretations.

We now introduce an entropic measure of common or mutual information shared by two parties. Suppose Alice possesses the random variable $A$ and Bob has the random variable $B$, the \textit{mutual information} specifies the number of bits in common between the two distributions. Equivalently, this represents the maximum number of bits that one party can learn about the other just by inspecting their own information.

For two classical distributions, the classical mutual information is given by,
\begin{align}\index{Mutual information}
I(A;B) = H(A) + H(B) - H(A,B).
\end{align}
Analogously, the quantum mutual information\index{Quantum mutual information} for bipartite state $\hat\rho_{A,B}$ is defined as,
\begin{align}
I(A;B)_{\hat\rho} = S(\hat\rho_A) + S(\hat\rho_B) - S(\hat\rho_{A,B}),
\end{align}
using the von Neuman entropy.

The mutual information between two quantum states is invariant under local unitary transformations,
\begin{align}
I(A;B)_{\hat\rho} = I(\hat{U}_A\hat\rho_A \hat{U}_A^\dag; \hat{U}_B\hat\rho_B \hat{U}_B^\dag),
\end{align}
since the eigenvalue spectrum of a density operator is invariant under unitary transformations. Therefore, the mutual information represents the maximum amount of information Bob can learn about Alice's state under \textit{any} local operations.

A quantum process cannot increase the mutual information between two parties. This yields the \textit{data processing inequality}\index{Data processing inequality} that, for a sequence of channels \mbox{$X\to Y\to Z$},
\begin{align}\index{Data processing inequality}\label{eq:data_proc_ineq}
I(X:Z)&\leq I(X:Y), \nonumber \\
I(X:Z)&\leq I(Y:Z),
\end{align}
with equality if and only if the channel not specified in the identity on the right hand side (\mbox{$Y\to Z$} or \mbox{$X\to Y$} respectively) is unitary, i.e one of the links in the chain perfectly preserves information content. The progression is shown in Fig.~\ref{fig:data_proc_ineq}.

\begin{figure}[!htbp]
\includegraphics[clip=true, width=0.3\textwidth]{data_proc_ineq}
\captionspacefig \caption{\label{fig:data_proc_ineq}A sequence of events \mbox{$X\to Y\to Z$}. The data processing inequality states that the mutual information from beginning to end is upper-bounded by the mutual information between neighbouring stages, as per Eq.~(\ref{eq:data_proc_ineq}).}	
\end{figure}

The mutual information is defined as being between a particular known pair of states. Of course, in a quantum channel we have the flexibility to manipulate the input state and the measurement at the output. From this, we can then define the \textit{classical information of the channel}\index{Classical information of the channel}, $I_c(\hat\rho_{A,B})$ -- the maximum of the mutual information, optimised over \textit{all} possible measurement settings. Suppose $\hat\rho_{A,B}$ is shared between Alice and Bob, from which they would like to extract maximal correlations (i.e joint information). They measure using local measurement bases $\{\Lambda_A^x\}$ and $\{\Lambda_B^x\}$, yielding random variables $A$ and $B$. Their quantum mutual information of the channel is defined as,
\begin{align}
I_c(\hat\rho_{A,B}) = \max_{\rho,\{\Lambda_A^x\},\{\Lambda_B^x\}} I(A;B). \label{eq:quant_mut}
\end{align}

The intuitive interpretation is that this is the maximum mutual information between input and output states that can be achieved across the channel. This effectively places a physical upper bound on the achievable bitrate or bandwidth of the channel.

%\subparagraph{Classical channel capacity}

The \textit{classical channel capacity} of $\mathcal{E}$ is defined as
\begin{align}\index{Classical!Channels!Capacity}
\mathcal{C}(\mathcal{E}) = \lim_{n \rightarrow \infty} \frac{I(\mathcal{E}^{\otimes n})}{n},
\end{align}
which can be interpreted as the maximum bitrate of the channel, per use of the channel, in the limit of an infinite number of uses, where entangled between multiple applications is allowed in general.

 For a noisy channel, if the input contains product states as well as separable measurements, the classical capacity is equal to Eq.~(\ref{eq:quant_mut}), that is, the single-use and the asymptotic classical capacities are equal.

%\subparagraph{}

If we also allow joint measurement (but not entanglement between input states), the classical channel capacity can be expressed as,
\begin{align}
\chi(\mathcal{E}) = \max_{p_i,\hat\rho_i} S(\hat\rho) - \sum_i p_i S(\hat\rho_i), \label{eq:holevoinfo}
\end{align}
which is known as the Holevo bound\index{Holevo bound}.

However, if entangled input states are allowed in conjunction with joint measurement setting, the Holevo bound no longer holds, and the asymptotic classical channel capacity,
\begin{align}
\chi(\mathcal{E}) < \mathcal{C(E)}.
\end{align}

It has been shown that using entangled input states in conjunction with joint measurements can increase the amount of classical information which can be transmitted over a noisy quantum channel. In this case, 
\begin{align}
C_\text{ent} = \lim _{n\to \infty} \frac{1}{n}\chi(\mathcal{E}^{\otimes n}).
\label{eq:ent_ent}
\end{align}

%\subparagraph{}

The private classical capacity $\mathcal{P(E)}$ of a quantum channel $\mathcal{E}$ describes the maximum rate at which the channel is able to send classical information through the channel reliably and privately, that is, the trusted parties do not want the environment to have access to their classical correlations.

Here Alice prepares the quantum state $\rho_x^A$ according to the probability distribution $\{ p_X(x) \}$. The expected density operator of the ensemble is of the form,
\begin{align}
\rho^{XA} \equiv \sum_x p_X(x)\ket{x}\bra{x}  \otimes p_x^A.
\end{align}

The difference between the amount of classical correlation that she can establish with Bob, minus the correlation obtainable by Eve is given by,
\begin{align}
I(X;B)_\rho - I(X;E)_\rho,
\end{align}
which leads to the definition of the \textit{private information of a quantum channel}, where Alice maximises over all of her inputs,
\begin{align}
I_\mathcal{P} = \max_{\rho^{XA}} \{I(X;B)_\rho - I(X;E)_\rho   \}.
\end{align}

The private classical capacity is defined as,
\begin{align}
\mathcal{P(E)} = \lim_{n\rightarrow \infty} \frac{1}{n} I_\mathcal{P}(\mathcal{E}^{\otimes n })
\end{align}

%\subparagraph{}
The last capacity in regards to classical communication over quantum channel is the \textit{entanglement-assisted classical capacity}, $\mathcal{C}_E(\mathcal{E})$. This quantity measures the classical information that can be transmitted through the channel, if Alice and Bob possess shared entanglement prior to the transmission,
\begin{align}
\mathcal{C}_E(\mathcal{E}) = \max_{p_i,\hat\rho_i} I(A:B).
\end{align}

The difference between $\mathcal{C(E)}$ and $\mathcal{C}_E(\mathcal{E}) $ is that $\mathcal{C}_E(\mathcal{E}) $ is equal to the maximised quantum mutual information (i.e additivity holds), and is equal to the single-use version of $\mathcal{C}_E(\mathcal{E}) $. This implies that shared entanglement does not change the additivity of quantum mutual information.

\paragraph{Quantum channel capacities}

The \textit{quantum channel capacity}\index{Quantum channel capacity} is then analogously defined as,
\begin{align}\label{eq:quantcohinfo}
Q(\mathcal{E}) = \lim_{n \rightarrow \infty} \frac{\mathcal{I}_Q(\mathcal{E}^{\otimes n})}{n}.
\end{align}

The most important distinction between quantum mutual information and quantum coherent information is that the mutual information is always additive, but the quantum coherent information is not.

Analytic solutions to Eq.~(\ref{eq:quantcohinfo}) are notoriously difficult to calculate. However, once net dephasing or depolarisation rates have been calculated across a route, the \textit{single-use} quantities can be calculated, which can serve as a lower bound to the ultimately achievable rates, if a number representing \textit{cost}, rather than an analytic solution, is all we need.

A summary of the different measure of classical and quantum capacities are given in Tab.~\ref{tab:capacities}.

\startnormtable
\begin{table*}[!htbp]
\begin{tabular}{ |c | c | c| c| } \hline
Capacity & Type of information & Correlation measure &Capacity formula  \\ \hline
Classical &Classical information &Holevo information &Eq.~(\ref{eq:holevoinfo})       \\
Private & Classical information & Private information  & See \cite{bib:PhysRevLett.103.120501} \\
Entanglement Assisted Classical & Classical information & Quantum mutual information & See \cite{bib:PhysRevLett.83.3081} \\
Quantum &Quantum information &Quantum Coherent Information &Eq.~(\ref{eq:quantcohinfo}) \\ \hline
\end{tabular}
\captionspacetab
\caption{\label{tab:capacities} Measure of classical and quantum channel capacities. Taken from \cite{bib:8242350}.}
\end{table*}
\startalgtable

%\paragraph{}

We now discuss the quantum capacity for the erasure channel and the amplitude damping channel, both of which the quantum capacities can be calculated.

We recall that the erasure channel acts on a state as follows is,
\begin{align}
\hat\rho \rightarrow (1-\eta) \hat\rho + \eta\ket{e}\bra{e}
\end{align}
where $\eta$ is the probability of losing the state and $\ket{e}\bra{e}$ is orthogonal to $\rho$. Given a system of dimension $d_A$, For $\eta \in [1/2,1]$, the quantum capacity is zero. For $\eta \in [0,1/2]$, the quantum capacity is $(1-2\eta)\log d_A.$

For an amplitude damping channel with damping parameter $\gamma$, the Kraus operators can be written as,
\begin{align}
A_0 &= \ket{0}\bra{0} + \sqrt{1-\gamma}\ket{1}\bra{1},\nonumber\\
A_1 &= \sqrt{\gamma}\ket{1}\bra{1}.
\end{align}

For $\gamma \in [0,1/2]$,the quantum capacity is equal to,
\begin{align}
\max_{p \in [0,1]} h_2 [(1-\gamma)p] - h_2[\gamma_p],
\end{align}
where $h_2$ represents the binary entropy function. For $\gamma \in [1/2,1]$, the quantum capacity is equal to zero.

%\paragraph{ }

Two of the most surprising discoveries of quantum Shannon theory is the \textit{superadditivity} of coherent information for the depolarising channel, and \textit{superactivation} of quantum capacity, where two channels whose individual capacities are zero can be combined to make a channel with non-zero capacity.

\subparagraph{Superadditivity of coherent information}

For qubits, the depolarising channel transmits its input with probability $(1-p)$ and replaces it with the maximally mixed state $\hat\openone/2$ with probability $p$,
\begin{align}
\rho \rightarrow (1-p) \rho + p \frac{\openone}{2}.
\end{align}
It is known that its coherent information is strictly superadditive when $p$ is large.

\begin{align}
5 I_\mathcal{Q}(\mathcal{E}) < 5 I_\mathcal{Q}(\mathcal{E}^{\otimes 5}).
\end{align}

It can be shown that the state which maximises the coherent information is the Bell state $\frac{1}{\sqrt2}(\ket{00}_{AA_1} + \ket{11}_{AA_1} )$, and the one-use coherent information is,
\begin{align} \label{eq:qcohdepol}
I_\mathcal{Q} = 1+ \left( 1-\frac{3p}{4} \right )\log\left[1-\frac{3p}{4}\right] + \frac{3p}{4}\log\left[\frac{p}{4}\right].
\end{align}

At $p = 0.25$, Eq.~(\ref{eq:qcohdepol}) is equal to zero, but if we calculate the coherent information for a five-qubit repetition code,
\begin{align}
\frac{1}{\sqrt2}&(\ket{00}_{AA_1} + \ket{11}_{AA_1} ) \rightarrow \nonumber \\ 
\frac{1}{\sqrt2}&(\ket{000000}_{AA_1A_2 A_3 A_4 A_5} + \ket{111111}_{AA_1A_2 A_3 A_4 A_5}),
\end{align}
at $p$ just above $0.25$, the quantity of interest is positive \cite{bib:PhysRevA.57.830}. This demonstrates that the quantum capacity of the depolarising channel
is superadditive. Despite the fact that the channel looks simple, the best strategy for achieving the quantum capacity of the depolarising channel remains very poorly understood \cite{bib:wilde2013quantum}.

\subparagraph{Superactivation}\index{Superactivation}

Suppose that Alice is connected to Bob by two quantum channel $\mathcal{E}_1,\mathcal{E}_2$ both with zero quantum capacity. Intuitively one would expect that Alice would not be able to transmit quantum information over the channel $\mathcal{E}_1\otimes \mathcal{E}_2$. However, examples of two zero-capacity channels are known to have non-zero joint quantum capacity. The phenomenon is known as \textit{superactivation}.

One of the channels is a 50\% erasure channel (which we know to have zero quantum capacity from the non-cloning theorem), and the other is known as an entanglement-binding channel \cite{bib:horodecki2001separability}.

This phenomenon has important implications for quantum information transmission. It implies that if there are other seemingly useless channels available, one could transmit more quantum information than if the channels were used alone. It also implies that the coherent information is strongly non-additive, and there is much to be understood about reliable communication rates over quantum channels.

@@

Def.~\ref{def:metric} defines the properties of a cost metric in the classical context. We now wish to consider this in the quantum context.

>>>>

The fidelity of two states directly quantifies how close they are to one another in a geometric sense, i.e on the Bloch sphere, or, in the context of a state passing through a quantum channel, a measure of how well the state is preserved.

@@

@@

Def.~\ref{def:metric} defines the properties of a cost metric in the classical context. We now wish to consider this in the quantum context, such that we are empowered to ask questions like ``what is the the total cost across a network route?'' or ``which route minimises a cost between two parties?'', where \textit{cost} now refers to some metric relevant to quantum state distribution, such as accumulated decoherence or loss.

### 2020-08-29 Sections/errors_in_quantum_networks.tex
editing Saturday night

A qubit dephasing channel is often quoted in terms of its $T_2$-time, a characteristic time for dephasing to occur under continuous time-evolution. Specifically, the probability of no dephasing occurring scales as,
\begin{align}
p_\mathrm{no\,error} = e^{-t/T_2},	
\end{align}
yielding a the equivalent dephasing channel,
\begin{align}
\mathcal{E}_t^\mathrm{dephasing}(\hat\rho) = e^{-t/T_2} \hat\rho + \frac{1}{2}(1-e^{-t/T_2})(\hat\rho + \hat{Z}\hat\rho\hat{Z}),
\end{align}
as shown in Fig.~\ref{fig:T2_time}. For a qubit density matrix,
\begin{align}
	\hat\rho = \begin{pmatrix}
 \alpha & \gamma \\
 \gamma^* & \beta
\end{pmatrix},
\end{align}
this is equivalent to adding a factor of $e^{-t/T_2}$ to the two off-diagonal (coherence) elements,
\begin{align}
	\hat\rho_t = \begin{pmatrix}
 \alpha & e^{-t/T_2}\gamma \\
 e^{-t/T_2}\gamma^* & \beta
\end{pmatrix}.
\end{align}

@@

\subsection{Depolarisation} \index{Depolarising channel}

>>>>

A qubit dephasing channel is often quoted in terms of its $T_2$-time, a characteristic time for dephasing to occur under continuous time-evolution. Specifically, the probability of no dephasing occurring scales as,
\begin{align}
p_\mathrm{no\,error} = e^{-t/T_2},	
\end{align}
yielding a the equivalent dephasing channel,
\begin{align} \label{eq:T2_time}
\mathcal{E}_t^\mathrm{dephasing}(\hat\rho) = e^{-t/T_2} \hat\rho + \frac{1}{2}(1-e^{-t/T_2})(\hat\rho + \hat{Z}\hat\rho\hat{Z}),
\end{align}
as shown in Fig.~\ref{fig:T2_time}. For a qubit density matrix,
\begin{align}
	\hat\rho = \begin{pmatrix}
 \alpha & \gamma \\
 \gamma^* & \beta
\end{pmatrix},
\end{align}
this is equivalent to adding a factor of $e^{-t/T_2}$ to the two off-diagonal (coherence) elements,
\begin{align}
	\hat\rho_t = \begin{pmatrix}
 \alpha & e^{-t/T_2}\gamma \\
 e^{-t/T_2}\gamma^* & \beta
\end{pmatrix}.
\end{align}

Note that Eq.~\eqref{eq:T2_time} is parameterised into an ideal term ($\hat\rho$) and a completely dephased term ($(\hat\rho+\hat{Z}\hat\rho\hat{Z})/2$), and thus multiple rounds of this channel yields an equivalent channel where the probability associated with the former term accumulates multiplicatively,
\begin{align}
e^{-t'/T_2} = \prod_i e^{-t_i/T_2}	
\end{align}
or in logarithmic form additively, making it applicable to cost vector analysis (Sec.~\ref{sec:cost_as_dist}),
\begin{align}
t' = \sum_i t_i,	
\end{align}
providing a direct mechanism for calculating the effective dephasing rate across multiple subsequent sections in a qubit network route. The same can easily be seen to apply to depolarising (Sec.~\ref{sec:depolarising_channel}), amplitude damping (Sec.~\ref{sec:amp_damp} and loss channels (Sec.~\ref{sec:eff_err}).

@@

\subsection{Depolarisation} \index{Depolarising channel} \label{sec:depolarising_channel}
